---
title: "Pipes"
execute: 
  echo: True
author: Cath Riddoch
format:
  html:
    embed-resources: true
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
library(tidyverse)
```

```{r}
head(mtcars)
```

no rstudio https://webr.r-wasm.org/latest/

## What are pipes?

%>%

|>


Keyboard shortcut 


## Example of using with or without pipe
Find all cars with 4 cylinders (cyl)
```{r}
filter(.data, condition)
```

```{r}

```

## Place holders

magrittr placeholder is .

base r placeholder is _

```{r}
mtcars %>%
  lm(mpg ~ wt, data = .)
```

```{r}
mtcars |>
  lm(mpg ~ wt, data = _)
```


## Example 2
```{r}
## find the mean mpg of cars with 4 cylinders

# get the cars with 4 cyl
four_cylinder_cars<- filter(mtcars, cyl == 4)

# get the mpg for those 4 cylinder cars
mpg_values<- four_cylinder_cars$mpg

# calculate the mean
result<- mean(mpg_values)

result
```

```{r}

```

```{r}

```










Or have nested function calls (don't do this)

```{r}
mean(filter(mtcars, cyl == 4)$mpg)
```


## Why use base pipe?


## Exercises
https://github.com/c-riddoch/training

1. Find the mass in kg of the heaviest 6 cylinder car (use wt_kg = wt * 453.6)

2. Return the name of the 6 cylinder car with highest mpg 

